---
title: "Gov 50 Final Project"
author: "Hazel Koh"
description: "Are democracies more likely to comply with International Court of Justice decisions on legal disputes?"
output:
  distill::distill_article:
    self_contained: false
---

## Introduction

Do democracies tend to comply more with International Court of Justice (ICJ) decisions regarding legal disputes? While ICJ decisions are legally binding, they are not enforceable. A state has full discretion over whether it will follow the recommendations of the Court. I hypothesize that democracies have a higher record of compliance than their counterparts, because democracies tend to have a higher respect for justice and the rule of law. I would find support for my hypothesis if I observe a statistically significant p-value when I run my regression of the compliance rate on the democracy variable, controlling for the judgement’s year of release and countries’ exports in current US$ as potential confounders. If there is no statistically significant relationship between the two main variables of democracy level and compliance rate, I will have to re-evaluate my hypothesis. 

The results would importantly shed light on the potential for international law to effectively facilitate international cooperation. A finding that democratic states have a higher respect for international law would reveal the shortcomings of international law in suggesting that complying with ICJ decisions remains more of a democratic norm than a global one. Moreover, a result of similarly low compliance rates by both democracies and non-democracies may suggest that international law does not yet play a significant role in state-to-state relations, while a finding of similarly high compliance rates would affirm the potential for international law to further cooperation among states regardless of their regime type.


## Data

To explore my hypothesis, I will be looking at contentious cases the ICJ has deliberated on between the years of 1947 and 2018. The main explanatory variable of interest is a state’s `Time-Varying Polity Score` in the year the judgement was released. The variable is measured through expert coding by the Center for Systemic Peace with a score of 10 for full democracies, scores of 6 to 9 for democracies, scores of 1 to 5 for open anocracies, scores of -5 to 0 for closed anocracies, and scores of -10 to -5 for autocracies. For some analyses, I will use this categorization of regime types, `level_of_dem`, to better illustrate which countries can be considered more or less democratic.  

My dependent variable is `Compliance` and is sourced from Dr. Tom Ginsburg’s data set for his book, “Democracies and International Law.” I conducted additional research on whether there was any progress in states complying with a case since Dr. Ginsburg released the data set and updated it accordingly. Cases are coded as 0 if states with the burden of compliance failed to comply and 1 if they followed the ICJ’s  recommendations. For the purpose of making a linear regression, I will use this binary `Compliance` variable to create a continuous average compliance variable named `comp_mean`.

To create a better logistic regression model of the binary compliance variable on states’ level of democracy, I will also be using the explanatory variables of `Judgement Year` gathered from Dr. Tom Ginsburg’s data set and exports of goods and services in current US$ using data from the World Bank. 

Despite using time series data that focuses on countries’ level of democracy over the years 1947-2018, I will be assuming that my research design is cross-sectional for the purpose of simplicity and staying within the scope of this class. With `Time-Varying Polity Score` as my treatment variable, I will be observing whether the treatment of being more democratic at the time of the judgement has an effect on compliance.  

```{r, include = FALSE}
library(tidyverse)
library(ggplot2)
icj_comp <- read_csv("icjcompliance2.csv")
icj_comp


icj_comp_bin <- icj_comp |>
  drop_na() |>
    mutate(
    level_of_dem = case_when(
      `Time-Varying Polity Score` == 10 ~ "Full Democracy",
      `Time-Varying Polity Score` < 10 & `Time-Varying Polity Score` >= 6 ~ "Democracy",
      `Time-Varying Polity Score` < 6 & `Time-Varying Polity Score` >= 1 ~ "Open Anocracy",
      `Time-Varying Polity Score` < 1 & `Time-Varying Polity Score` >= -5 ~ "Closed Anocracy",
      `Time-Varying Polity Score` < -5 & `Time-Varying Polity Score` >= -10 ~ "Autocracy"),
    compliance = if_else(Compliance == 1, "Compliance", "Non-compliance"),
    level_of_dem = fct_relevel(as.factor(level_of_dem), levels = c("Full Democracy", "Democracy", "Open Anocracy", "Closed Anocracy", "Autocracy"))) |>
rename(exports = `Exports of Goods and Services (Current $USD)`)


```
##### Visual Summary of Dependent `Compliance` Variable:
```{r Visual Summary of Dependent Variable, echo = FALSE}
dem_comp_plot <- icj_comp_bin |>
  ggplot(aes(x = as.factor(level_of_dem),
             fill = as.factor(level_of_dem))) +
  geom_bar(binwidth = 1) +
  labs(x = "",
       y = "Count") +
  facet_wrap(~ compliance) +
  theme(legend.position = "none",
        axis.text.x = element_text(size = 5)) +
  labs(x = "Regime Types") +
  theme(plot.caption = element_text(hjust = 0.5))
dem_comp_plot
```
In this visualization of the dependent `Compliance` variable, we can observe some differences in the distributions of regime types across two groups with different outcome values: one with compliance, and the other with non-compliance. For example, the compliance outcome reveals there are more cases of democracies complying than full democracies, autocracies, and open anocracies. Among the less democratic regime types, autocracies with the lowest Polity scores actually have the highest record of compliance. Moreover, the non-compliance outcome illustrates that there are overall fewer cases of non-compliance than compliance. Among these cases, autocracies have the lowest record of noncompliance, while full democracies have the highest. 


## Results and Analysis
### Linear Regression
##### Visualization of Average Compliance with ICJ Decisions per Polity Score:
```{r , include = FALSE }
library(tidyverse)
library(ggplot2)
icj_comp <- read_csv("icjcompliance2.csv")
icj_comp

comp_pol_mean <- icj_comp |>
  drop_na() |>
    mutate(
    exports = `Exports of Goods and Services (Current $USD)`/1000000000,
    level_of_dem = case_when(
      `Time-Varying Polity Score` == 10 ~ "Full Democracy",
      `Time-Varying Polity Score` < 10 & `Time-Varying Polity Score` >= 6 ~ "Democracy",
      `Time-Varying Polity Score` < 6 & `Time-Varying Polity Score` >= 1 ~ "Open Anocracy",
      `Time-Varying Polity Score` < 1 & `Time-Varying Polity Score` >= -5 ~ "Closed Anocracy",
      `Time-Varying Polity Score` < -5 & `Time-Varying Polity Score` >= -10 ~ "Autocracy"),
   level_of_dem = fct_relevel(as.factor(level_of_dem), levels = c("Full Democracy", "Democracy", "Open Anocracy", "Closed Anocracy", "Autocracy"))) |>
  group_by(level_of_dem) |>
  summarize(comp_mean =  mean(Compliance, na.rm = TRUE)) 
```
```{r, echo = FALSE}
comp_mean <- icj_comp |>
  drop_na() |>
  group_by(`Time-Varying Polity Score`) |>
  summarize(comp_mean =  mean(Compliance, na.rm = TRUE)) 

comp_mean_plot <- comp_mean |>
  ggplot(aes(x = `Time-Varying Polity Score`,
         y = comp_mean)) +
  geom_col(mapping = aes(fill = factor(`Time-Varying Polity Score`))) +
  theme_minimal() +
  labs(x = "Polity Score",
       y = "Average Compliance") +
  theme(legend.position = "none")
comp_mean_plot
```

##### Visualization of Average Compliance with ICJ Decisions per Regime Type:
```{r Data Visualization: Polity Score Breakdown, echo = FALSE}
comp_pol_plot <- comp_pol_mean |>
  ggplot(aes(x = level_of_dem,
         y = comp_mean,
         fill = level_of_dem)) +
  geom_col() +
  theme_minimal() +
  labs(x = "Regime Types",
       y = "Average Compliance") +
  theme(legend.position = "none")
comp_pol_plot
```


```{r, include = FALSE}
icj_comp_bin <- icj_comp |>
  drop_na() |>
    mutate(
    exports = `Exports of Goods and Services (Current $USD)`/1000000000,
    level_of_dem = case_when(
      `Time-Varying Polity Score` == 10 ~ "Full Democracy",
      `Time-Varying Polity Score` < 10 & `Time-Varying Polity Score` >= 6 ~ "Democracy",
      `Time-Varying Polity Score` < 6 & `Time-Varying Polity Score` >= 1 ~ "Open Anocracy",
      `Time-Varying Polity Score` < 1 & `Time-Varying Polity Score` >= -5 ~ "Closed Anocracy",
      `Time-Varying Polity Score` < -5 & `Time-Varying Polity Score` >= -10 ~ "Autocracy"),
    compliance = if_else(Compliance == 1, "Compliance", "Non-compliance"),
    level_of_dem = fct_relevel(as.factor(level_of_dem), levels = c("Full Democracy", "Democracy", "Open Anocracy", "Closed Anocracy", "Autocracy"))) 
```


```{r, include = FALSE}
library(broom)
lin_reg <- icj_comp |>
  drop_na() |>
  group_by(`Time-Varying Polity Score`) |>
  summarize(comp_mean =  mean(Compliance, na.rm = TRUE))

fit1 <- lm(comp_mean ~ `Time-Varying Polity Score`, data = lin_reg)
fit1
```

```{r, include = FALSE}
fit1 |>
  broom::tidy() |>
  dplyr::select(term, estimate) |>
  knitr::kable(caption = "Coefficient Estimates for a Model Fitted to Estimate Variation in Compliance")
```

##### Coefficient Estimates for a Linear Model Fitted to Estimate Variation in Average Compliance:
```{r, echo = FALSE}
var_labels <- c(
  "(Intercept)" = "Intercept",
  "Time-Varying Polity Score" = "Time-Varying Polity Score"
)

modelsummary::modelsummary(
  fit1,
  coef_map = var_labels,
  statistic = c("s.e. = {std.error}",
                "p = {p.value}"),
  gof_map = c("nobs", "r.squared", "adj.r.squared", ""))
```

##### Visualization of Linear Regression Model:
```{r Visualization of Linear Regression, echo = FALSE}
ggplot(lin_reg, 
       aes(x=`Time-Varying Polity Score`, y=comp_mean)) +
  geom_point(color = "steelblue1") +
  geom_smooth(method="lm", se=FALSE) +
  labs(x = "Polity Score",
       y = "Average Compliance")
```
According to the linear regression of average compliance on countries’ Polity scores, countries with a Polity score of 0 have a 71.3% likelihood of complying with ICJ decisions. With every 1 unit increase in its Polity score, a country is estimated to be 0.16 percentage points less likely to comply with an ICJ decision. The two bar plots also illustrate that countries with lower Polity scores actually have a higher rate of compliance than countries with higher Polity scores. Since the estimated coefficient of `Time-Varying Polity Score` is not statistically significant with a p-value of 0.918, however, no statistically significant association between Polity score and likelihood of compliance with ICJ judgments can be proved. 

### Logistic Regression
```{r , include = FALSE}
fit2 <- glm(Compliance ~ `Time-Varying Polity Score`, data = icj_comp_bin, family = binomial())
fit2
summary(fit2)

fit2 |>
  broom::tidy() |>
  knitr::kable(caption = "Coefficient Estimates for a Model Fitted to Estimate Variation in Compliance", digits = c(0,2,3,2,3), col.names = c("Predictor", "Estimate", "Std. Error", "Statistic", "P-value"))

fit3 <- glm(Compliance ~ `Time-Varying Polity Score` + `Judgement Year` + exports, data = icj_comp_bin, family = binomial())
fit3
summary(fit3)


library(equatiomatic)
extract_eq(fit3)

fit3 |>
  tidy() |>
  knitr::kable(caption = "Coefficient Estimates for a Model Fitted to Estimate Variation in Compliance",
        col.names = c("Predictor", "Estimate", "Std. Error", "Statistic", "P-value"))
```
##### Coefficient Estimates for a Logistic Model Fitted to Estimate Variation in Compliance:
```{r , echo = FALSE}
var_labels2 <- c(
  "(Intercept)" = "Intercept",
  "Time-Varying Polity Score" = "Time-Varying Polity Score",
  "Judgement Year" = "Judgement Year",
  "exports" = "Exports in Current US$"
)

modelsummary::modelsummary(
  list(fit2, fit3),
  coef_map = var_labels2,
  statistic = c("s.e. = {std.error}",
                "p = {p.value}"),
  gof_map = c("nobs", "r.squared", "adj.r.squared", ""))
```


##### Visualization of Logistic Regression Model:
```{r Visualization of Logistic Regression, echo = FALSE}
library(sjPlot)
plot_model(fit3, type = "pred", terms = "Time-Varying Polity Score") +
  labs(x = "Polity Score",
       y = "Compliance",
       title = "Predicted Probabilities of Compliance per Polity Score")
```
A logistic regression of the binary variable of `Compliance` on a country’s Polity score controlling for the judgement year and a country’s exports in current USD$ shows that a 1 unit increase in a country’s Polity score is associated with a 0.013 unit decrease in the odds of compliance. However, with a p-value of 0.837, the estimated coefficient of `Time-Varying Polity Score` is also not statistically significant in this logistic regression. Therefore, we cannot assume that there is a statistically significant association between Polity score and likelihood of compliance with ICJ judgments. 


## Conclusion
With no evidence of a statistically significant association between a country’s Polity score and compliance, the hypothesis that democracies are more likely to comply with ICJ judgments cannot be proved and must either be rejected or re-evaluated. 

Some limitations of this research design include the difficulty of measuring compliance, the data’s outdatedness, missing values, and unaccounted for confounders. Since there is no date by which states are required to comply with a decision, it may be that states who were coded to have failed to comply thus far intend to comply in the future. Moreover, the accuracy of the compliance values may be undermined due to a lack of information. For example, in cases involving reparations, it may be difficult to measure a state’s compliance if neither parties publicize whether reparations were made. It would be useful to have access to such information for more accurate results. Another limitation is that the data set is from 2018 and lacks judgments that have been made since. To address this limitation, the current data set should be updated with recent ICJ cases to increase the accuracy and representativeness of the sample. Increasing the sample of cases may also help to improve the results and accordingly its external validity. Similarly, cases in which judgments were made post-2015 lack Polity scores and cases with judgments made before 1960 have missing data on exports. It would have been useful to have had access to more recent Polity scores and past exports data. After dropping missing values, I was left with a far smaller sample size of 38 cases as opposed to 146 cases. To improve internal validity, it may also be helpful to address different confounders that I could not control for because they were not in the data set, such as GDP per capita and measures of military power.

To better answer my research question about whether there is an association between democracies and the likelihood of complying with ICJ decisions, therefore, I will need to implement these changes before re-evaluating my hypothesis. 
